CFF GCT Study

### 2025-01-24 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
updates for Monique (GCT)

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(gtsummary)
library(gt)
library(rspiro)
library(childsds)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
```

@@

@@

      Type.2.diabetes.in.the.family...Father,

@@

      Type.2.diabetes.in.the.family...Mother,
    "Brothers (number with diabetes)" = Brothers..number.with.diabetes.,
    "Sisters (number with diabetes)" = Sisters..number.with.diabetes.,

@@

@@

```{r}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
df %>%
  select(-ID) %>%
  tbl_summary(
    by = `Got OGTT and GCT?`, missing_text = "Missing",
    type =
      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
        "continuous2"
  ) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma or OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

```{r}
#| label: table1b
#| tbl-cap: Participant characteristics by CFRD status
#| message: false
df %>%
  select(-ID) %>%
  tbl_summary(
    by = Diabetes, missing_text = "Missing",
    type =
      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
        "continuous2"
  ) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: table1c
#| tbl-cap: Participant characteristics by dysglycemia
#| message: false
df %>%
  select(-ID) %>%
  tbl_summary(
    by = Dysglycemia, missing_text = "Missing",
    type =
      "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" ~
        "continuous2"
  ) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

@@

```{r}
invisible(lapply(
  list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
  function(roc) {
    print(plot(roc, print.auc = T, print.thres = T))
  }
))
```

>>>>

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
knitr::opts_knit$set(root.dir = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF")
```

@@

  mutate(

@@

      factor(Type.2.diabetes.in.the.family...Father),

@@

      factor(Type.2.diabetes.in.the.family...Mother),
    "Brothers (number with diabetes)" = factor(Brothers..number.with.diabetes.),
    "Sisters (number with diabetes)" = factor(Sisters..number.with.diabetes.),

@@

# Drop levels
df$Race <- droplevels(df$Race)
# Set table options
mycontrols <-
  tableby.control(
    numeric.stats = c("N", "Nmiss2", "meansd", "range"),
    cat.stats = c("N", "Nmiss2", "countpct")

@@

## By OGTT status

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma or OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

## By CFRD status

```{r results='asis'}
#| label: table1b
#| tbl-cap: Participant characteristics by CFRD status
#| message: false
t1b <- tableby(update(f, Diabetes ~ . - Diabetes),
  data = df, control = mycontrols
)
summary(t1b, pfootnote = T, digits = 2)
```

### Pairwise comparisons

```{r}
pairwise.t.test(df$`Average of Highest FEV1pp`, df$Diabetes)
pairwise.fisher.test(df$`Pancreatic insufficient`, df$Diabetes,
  p.adjust.method = "holm"
)
pairwise.t.test(df$HbA1c, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min plasma`, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min capillary`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 0 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 60 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 120 min`, df$Diabetes)
pairwise.wilcox.test(df$`Days From OGTT to GCT`, df$Diabetes)
```

## By dysglycemia status

```{r results='asis'}
#| label: table1c
#| tbl-cap: Participant characteristics by dysglycemia
#| message: false
t1c <- tableby(update(f, Dysglycemia ~ . - Dysglycemia),
  data = df, control = mycontrols
)
summary(t1c, pfootnote = T, digits = 2)
```

@@

```{r}
invisible(lapply(
  list(gctpl_roc, gctcap_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
  function(roc) {
    print(plot(roc, print.auc = T, print.thres = T, legacy.axes = T))
  }
))
```
